---
title: "5a_wdi_countries"
author: "Leslie Roberson"
date: "3/7/2020"
output: html_document
---

## Summary

Here I download WGI data and prepare it for the linear regression for threatened species catch and imports

# Data source

reference: http://info.worldbank.org/governance/wgi/index.aspx#home
downloaded: 30 Sep 2019

# Description:

The Worldwide Governance Indicators (WGI) project reports aggregate and individual governance indicators for 215 economies over the period 1996–2017, for six dimensions of governance:

Voice and Accountability
Political Stability and Absence of Violence
Government Effectiveness
Regulatory Quality
Rule of Law
Control of Corruption

Time range: 1996-2017

# GDP and GDP per capita

pull GDP (current USD) and GDP per capita (current USD) from the WDI database

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(WDI) # install.packages("WDI")
library(wbstats) # install.packages("wbstats")
library(stringr)

```

## specify time frame

```{r time frame}

# pull 10 years bc I'll select the most recent year in this timeframe for each country
yr_start = 2005
yr_end   = 2015

```

## Download each of the 6 WGI indicators

```{r download WGI indicators}

## get description of variables:
indicators <-  data.frame(WDI_data[[1]])
indicators[grep("VA.EST", indicators$indicator), ]
indicators[grep("PV.EST", indicators$indicator), ]
indicators[grep("GE.EST", indicators$indicator), ]
indicators[grep("RQ.EST", indicators$indicator), ]
indicators[grep("RL.EST", indicators$indicator), ]
indicators[grep("CC.EST", indicators$indicator), ]

# identify the six indicators
# WDIsearch('violence')# general search
key_voice = WDI(
  indicator = WDIsearch('Voice and Accountability: Estimate', field='name')['indicator'],
  country = 'all', start = yr_start, end=yr_end)
key_voice <- key_voice %>% rename(voice = indicator)

key_polst = WDI(
  WDIsearch('Political Stability and Absence of Violence/Terrorism: Estimate', field='name')['indicator'],
  country='all',start = yr_start, end=yr_end)
key_polst <- key_polst %>% rename(polst = indicator)

key_gvtef = WDI(
  WDIsearch('Government Effectiveness: Estimate', field='name')['indicator'],
  country='all',start = yr_start, end=yr_end)
key_gvtef <- key_gvtef %>% rename(gvtef = indicator)

key_regqt = WDI(
  WDIsearch('Regulatory Quality: Estimate', field='name')['indicator'],
  country='all',start = yr_start, end=yr_end)
key_regqt <- key_regqt %>% rename(regqt = indicator)

key_rolaw = WDI(
  WDIsearch('Rule of Law: Estimate', field='name')['indicator'],
  country='all',start = yr_start, end=yr_end)
key_rolaw <- key_rolaw %>% rename(rolaw = indicator)

key_corrp = WDI(
  WDIsearch('Control of Corruption: Estimate', field='name')['indicator'],
  country='all',start = yr_start, end=yr_end)
key_corrp <- key_corrp %>% rename(corrp = indicator)

```

Combine the indicators into a single table, with a column for each indicator, and rows for each country-year pair

```{r indicators combined table}

d <- key_voice %>% 
  select(country, iso2c, year, voice) %>%
  left_join(key_polst, by=(c('country', 'iso2c','year'))) %>%
  left_join(key_gvtef, by=(c('country', 'iso2c','year'))) %>%
  left_join(key_regqt, by=(c('country', 'iso2c','year'))) %>%
  left_join(key_rolaw, by=(c('country', 'iso2c','year'))) %>%
  left_join(key_corrp, by=(c('country', 'iso2c','year'))); head(d); summary(d); sapply(d, class) 

names(d)
# check for missing years
summary(d)

```

## Determine which countries have missing years

```{r row has na}

row.has.na <- apply(d, 1, function(x){any(is.na(x))})

sum(row.has.na) 

rows.na <- d[row.has.na,]

```

## filter dataframe for most recent year available for each country

```{r filter most recent year}

# For each country take most recent year that has all 6 indicators
d_max <- d %>% 
  drop_na() %>% 
  group_by(country) %>%
  top_n(1, year) %>% # take top 1 year for each country
  ungroup()

summary(d_max)

duplicated(d_max$country) 

```

## Calculate overall WGI score for each country

```{r cum WGI score}
d_calcs  <-  d_max %>%
  group_by(country, iso2c, year) %>%
  rowwise() %>%
  mutate(score_wgi_scale = mean(voice:corrp, na.rm=T)) %>%
  ungroup()

summary(d_calcs)

```

## Add the GDP data

Also from World Bank using the WDI() package

The most recent GDP data available for each country (reported in US dollars for that year) are accessed from the World Bank (data.worldbank.org/indicator/NY.GDP.MKTP.CD)

```{r get GDP from WDI()}

## Get GDP and GDP per capita from WDI

WDIsearch('gdp.*current')
# current GDP : "NY.GDP.MKTP.CD"      "GDP (current US$)" 
# current GDP per capita: "NY.GDP.PCAP.CD" (US$)

# GDP
gdp_usd <- WDI(indicator = "NY.GDP.MKTP.CD", country = 'all', start = 2005, end = 2015)
# GDP per capita
gdp_percap_usd <- WDI(indicator = "NY.GDP.PCAP.CD", country = 'all', start = 2005, end = 2015)

gdp_usd <- gdp_usd %>% rename(gdp_usd = NY.GDP.MKTP.CD) 
gdp_percap_usd <- gdp_percap_usd %>% rename(gdp_percap_usd = NY.GDP.PCAP.CD)

# check for NAs
gdp_usd %>% filter(is.na(gdp_usd)) 
gdp_percap_usd %>% filter(is.na(gdp_percap_usd)) 

# take the 2015 or most recent gdp val for each country
gdp_usd_max <- gdp_usd %>% 
  drop_na() %>% # in any column
  group_by(country) %>%
  top_n(1, year) %>% # take top 1 year for each country
  ungroup()

gdp_percap_usd_max <- gdp_percap_usd %>% 
  drop_na() %>% # in any column
  group_by(country) %>%
  top_n(1, year) %>% # take top 1 year for each country
  ungroup()

summary(gdp_max) 

gdp_usd_max <- gdp_usd_max %>% arrange(country) %>% unique()
gdp_percap_usd_max <- gdp_percap_usd_max %>% arrange(country) %>% unique()

names(gdp_usd_max)
names(gdp_percap_usd_max)

gdp_max <- full_join(gdp_usd_max, gdp_percap_usd_max, by = c("iso2c","country","year"))

```

## remove all the aggregated regional values

```{r remove regions}

wb_countries <- wbcountries()
names(wb_countries)

gdp_max <- merge(gdp_max, y = wb_countries[c("iso2c","region","country")], by = "iso2c", all.x = TRUE)

gdp_max <- filter(gdp_max, region != "Aggregates")

summary(gdp_max) # 210 countries

gdp_max <- gdp_max %>% select(-country.y) %>% rename(country = country.x, year_gdp = year)

```

## join GDP to WDI indicators

```{r join gdp to WDI}

names(d_calcs)
names(gdp_max)

wdi_dat <- full_join(d_calcs, gdp_max, by = c("iso2c", "country"))

summary(wdi_dat)

```

## Get total population data from WDI

```{r WDI population}

# search avaialble data
WDIsearch(string = "population", field = "name", short = TRUE, cache = NULL)

WDIsearch('population.*total')

# population
population <- WDI(indicator = "SP.POP.TOTL", country = 'all', start = 2005, end = 2015)
head(population)

population <- population %>% rename(pop = SP.POP.TOTL) 

population <- population %>% filter(country != "Not classified")
names(population)

population_max <- population %>% 
  drop_na() %>% # in any column
  group_by(country) %>%
  top_n(1, year) %>% # take top 1 year for each country
  ungroup()

names(population_max)
population_max <- population_max %>% rename(year_pop = year)
population_max %>% filter(is.na(pop)) 

```

## Add population data to the wdi data

```{add population}

names(wdi_dat)
names(population_max)

wdi_dat <- full_join(wdi_dat, population_max, by = c("iso2c", "country"))

wdi_dat %>% filter(is.na(pop)) # n = 9 countries

## save

write_csv(wdi_dat, "_data/countries_wdi.csv")

```

